---
title: "project_4_D"
author: "Daryna"
date: "12/30/2021"
output: html_document
---

According to BBC, streaming online platforms became more popular than ever during the lockdown provoked by the COVID-19 pandemic. For instance, in UK during April 2020, 40% of the waking hours were spent in front of the screens, on average. Because of the growing popularity of online streaming services all over the world and the increasing variety of content available on such platforms, in this report we propose to analyze 4 major platforms: Netflix, Prime Video, Hulu and Disney+ in terms of quality. We specifically concentrate on movies that are available on these platforms in the United States. Depending on the platform, the monthly price for subscription varies: in 2021, Netflix costs - 9$, Prime Video - 9$, Hulu - 7$ and Disney+ 8$. The aim of this report is to understand which platform leads in terms of quality content and suits better potential future users, for example depending on the genre preference or age group. 
The analysis is performed on the dataset available on Kaggle that originally was scraped on the streaming platforms. Specifically, the dataset contains the list of movies available on the 4 streaming pltaforms in 2021. Overall, the dataset contains 9515 row where each row corresponds to a movie and is characterized by the following variables: 
- ID, unique ID for each movie,
- Title, full title of the movie, 
- Year, release year of the movie, time interval from 1914 to 2021,
- Age, factor variable specifying age group of the audience for which the movie is intended with the following levels: "all", "7+", "13+","16+" and "18+",
- Rotten Tomatoes, score out of 100 that intends to measure the quality of movies provided by Rotten Tomatoes - an american review-aggregation website for film and television,
- IMBd, score out of 10 that intends to measure the quality of movies provided by IMBd which is an online database of information related to films and television series, including ratings.
- Netflix, binary variable taking 1 if the movie is available on Netflix streaming platform, 0 - otherwise,
- Prime Video, binary variable taking 1 if the movie is available on Prime Video streaming platform, 0 - otherwise,
- Hulu, binary variable taking 1 if the movie is available on Hulu streaming platform, 0 - otherwise,
- Disney+, binary variable taking 1 if the movie is available on Disney+ streaming platform, 0 - otherwise.


Reference:
https://www.bbc.com/news/entertainment-arts-53637305
https://www.cnet.com/tech/services-and-software/best-streaming-service-of-2022/
https://www.kaggle.com/ruchi798/movies-on-netflix-prime-video-hulu-and-disney

```{r}
# Load libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(plotly)
require(gridExtra)
library(scales)
library(stringr)
library(RColorBrewer)
require("ggrepel")
```

```{r}
# Load Dataset
data <- read.csv("~/Documents/GitHub/SCV_project4/MoviesOnStreamingPlatforms_updated.csv")
```

```{r}
# Pie chart of the proportions of the moviesin the dataset
ggplot(total_by_platform, aes(x = "" , y = V1, fill = fct_inorder(names))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel1") +
  geom_label_repel(data = total_by_platform2,
                   aes(y = pos, label = paste0(V1)),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Streaming platform")) + scale_fill_manual(values=c("#E50914", "#00A8E1", "#1CE783", "#006E99")) +
  labs(title = "Composition of the total available movies by platform") +
  theme_void()
```



```{r}
# Distribution of the rating for RT  
plot1 <- ggplot(data = subset(data, !is.na(RT_num)), aes(x=RT_num))+
  geom_histogram(color="darkblue", fill="#F93208", bins = 50) + 
  labs(x = "Rotten tomatoes score") + 
  geom_vline(aes(xintercept=mean(RT_num)),
            color="black", linetype="dashed", size=0.5) + theme_bw()


# Distribution of the rating for IMDb 
plot2 <- ggplot(data = subset(data, !is.na(IMDb_num)), aes(x=IMDb_num))+
  geom_histogram(color="darkblue", fill="#DBA506", bins = 50) + 
  labs(x = "IMDb score") +
  geom_vline(aes(xintercept=mean(IMDb_num)),
            color="black", linetype="dashed", size=0.5) + theme_bw()

grid.arrange(plot1, plot2, ncol=2, top= "Distribution of the ratings' scores")
```


```{r}
## GROUPED Boxplots
ggplot(by_platform_review, aes(x= Platform, y=Rating, fill = Review, color  = Platform)) + 
  scale_fill_manual(values=c("#DBA506", "#F93208")) + 
                      scale_color_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1")) + theme_bw() + labs(title = "Distribution of scores by platform") + 
  geom_boxplot()
```

```{r}
## Plot of time series
ggplot(by_year_platform, aes(x = Year, y = n)) + 
  geom_line(aes(color = Platform), size = 1) +  scale_color_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1")) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+ scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + labs(title = "Number of movies available by platform depending on the release year", y = "Count") +theme_bw()
```


```{r}
plot1 <- ggplot(subset(data, !is.na(RT_num)), aes(x = Year, y = RT_num)) + geom_point(alpha = 0.3, color =  "#F93208") + geom_smooth(method = "lm") + scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + labs(y = "Rotten Tomatoes score")+ theme_bw()

plot2 <- ggplot(subset(data, !is.na(IMDb_num)), aes(x = Year, y = IMDb_num)) + geom_point(alpha = 0.5, color = "#DBA506") + geom_smooth(method = "lm") + scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + labs(y = "IMDb score")  + theme_bw()


grid.arrange(plot1, plot2, ncol=2, top= "Rating scores over the years")
```

```{r}
ggplot(by_year_rating, aes(x =n, y = reorder(genres_categ, n), fill = platform)) + 
  geom_bar(stat="identity", color='black')+ scale_fill_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1"), name="Streaming platform") + labs(title = "Number of movies by genre depending on the platform", y = "Genre", x = "Number of movies") + theme_bw()
```


```{r}
## By platform
ggplot(subset(less_genres_data, Netflix == 1 & RT_num >60 ), aes(x = Year, y = RT_num, color = genres_categ)) + geom_point( alpha = 0.5, size = 1.5) + scale_color_brewer(palette = "Set1", name= "Genre") + labs(title = "Well rated Netflix movies by release year and genre", y = "Rotten Tomatoes score")+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_bw()
```
```{r}
ggplot(subset(less_genres_data, Hulu == 1 & RT_num >60 ), aes(x = Year, y = RT_num, color = genres_categ)) + geom_point( alpha = 0.5, size = 1.5) + scale_color_brewer(palette = "Set1", name ="Genre") + labs(title = "Well rated Hulu movies by release year and genre", y = "Rotten Tomatoes score")+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_bw()
```

```{r}
ggplot(subset(less_genres_data, Disney. == 1 & RT_num >60 ), aes(x = Year, y = RT_num, color = genres_categ)) + geom_point( alpha = 0.5, size = 1.5) + scale_color_brewer(palette = "Set1", name ="Genre") + labs(title = "Well rated Disney+ movies by release year and genre", y = "Rotten Tomatoes score")+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_bw()
```


```{r}
ggplot(subset(less_genres_data, Prime.Video == 1 & RT_num >60 ), aes(x = Year, y = RT_num, color = genres_categ)) + geom_point( alpha = 0.5, size = 1.5) + scale_color_brewer(palette = "Set1", name ="Genre") + labs(title = "Well rated Prime Video movies by release year and genre", y = "Rotten Tomatoes score")+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_bw()
```

```{r}
ggplot(top_rated, aes(x = Year, y = RT_num, color = genres_categ)) + geom_point( alpha = 0.5, size = 1) + 
  geom_text_repel(aes(label = rownames(top_rated)),
                  size = 2) +scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + scale_color_brewer(palette = "Set2",  name ="Genre") + labs(title = "Titles of the top rated movies by release year and genre", y = "Rotten Tomatoes score") +theme_bw()
```


```{r}
ggplot(top_rated, aes(x = Year, y = RT_num, color = platform)) + geom_point( alpha = 0.5, size = 1) + 
  geom_text_repel(aes(label = rownames(top_rated)),
                  size = 2) + scale_color_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1"),  name ="Streaming Platform") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Titles of the top rated movies by release year and platform", y = "Rotten Tomatoes score")  + theme_bw()

```


```{r}
data$platform<- as.character(data$platform)
data$platform[data$platform == "Prime.Video"] <-"Prime Video"
data$platform[data$platform == "Disney"] <-"Disney+"

ggplot(data, aes(x = Year, y = RT_num, color = platform)) + geom_point( alpha = 0.3, size = 1) + scale_color_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1"), name= "Streaming Platform") + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Rotten Tomatoes score by year and platform", y = "Rotten Tomatoes score") + theme_bw()
```

```{r}
ggplot(data,aes(Year, RT_num, color = platform)) +
  stat_summary(geom = "line", fun.y = mean) + scale_color_manual(values = c("#006E99", "#1CE783","#E50914", "#00A8E1"), name= "Streaming Platform") + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Average rotten Tomatoes score by release year and platform", y = "Score") + theme_bw()

```


```{r}
ggplot(quality_composition, aes(fill=RT_quality_categ, y=platform, x=n)) + 
  geom_bar(stat="identity") + scale_fill_brewer(palette = "YlOrRd", name ="Quality",  direction=-1) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Composition by quality of movies by platform (Rotten tomatoes)", y = "Platform", x = "Number of movies")+ theme_bw()

```


```{r}
#Percentage 
ggplot(quality_composition, aes(fill=RT_quality_categ, y=platform, x=n)) + 
  geom_bar(position="fill", stat="identity") +  scale_fill_brewer(palette = "YlOrRd", name ="Quality",  direction=-1) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +  labs(title = "Composition by quality of movies by platform (Rotten tomatoes)", y = "Platform", x = "Proportion of movies") + theme_bw()
```


```{r}
#Genres in top movies Genres
ggplot(genres_composition, aes(fill=genres_categ, y=platform, x=n)) + 
  geom_bar(stat="identity") + scale_fill_brewer(palette = "Set3", name ="Quality",  direction=-1) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Composition by genres of top rated movies by platform", subtitle = "Rotten tomatoes, rating 80/100and higher", y = "Platform", x = "Number of movies")+ theme_bw()

```

```{r}
ggplot(age_composition, aes(fill=Age, y=platform, x=n)) + 
  geom_bar(stat="identity") + scale_fill_brewer(palette = "Set3", name ="Quality",  direction=-1) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + labs(title = "Composition by age groups of top rated movies by platform",subtitle = "Based on rotten tomatoes score, rating 80/100 and higher", y = "Platform", x = "Number of movies")+ theme_bw()
```

